---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r, cache = TRUE}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip")
unzip("activity.zip")
```

```{r}
getwd()
data_withna = read.table("activity.csv", header = TRUE, sep = ',')
data = data_withna[complete.cases(data_withna),]
head(data)
names(data)
summary(data)
```

## What is mean total number of steps taken per day?
**Calculate the total number of steps taken per day**
```{r}
total_steps_per_day = aggregate(data$steps, 
                                by = list(data$date), 
                                sum)$x
total_steps_per_day
```
**Make a histogram of the total number of steps taken each day**
```{r qplot, fig.width=8, fig.height=6, message=FALSE}
library(ggplot2)
hist(total_steps_per_day, 
     main = "Histogram of Steps/Day", 
     ylab = "Frequency", 
     xlab = "Steps/day")
```
**Calculate and report the mean and median of the total number of steps taken per day**
**mean**
```{r}
mean(total_steps_per_day)
```
**median**
```{r}
median(total_steps_per_day)
```
## What is the average daily activity pattern?
**Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**
```{r}
average_daily_activity_pattern = aggregate(data$steps, by = list(data$interval), mean)
head(average_daily_activity_pattern)
with(average_daily_activity_pattern, 
     plot(Group.1, 
          x, 
          type='l', 
          main = "Average daily activity pattern", 
          xlab = "5 min time interval", 
          ylab = "average daily activity"))

```

**Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**
```{r}
average_daily_activity_pattern[which.max(average_daily_activity_pattern$x),]
```

## Imputing missing values
**Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)**
```{r}
sapply(data_withna, function(x) sum(is.na(x)))
```
**Devise a strategy for filling in all of the missing values in the dataset.**
***The Strategy we are adopting is to replace na of any day with the mean steps taken for that day***

**Create a new dataset that is equal to the original dataset but with the missing data filled in.**
```{r}
for(date in unique(data_withna$date))
{
    if(any(is.na(data_withna[data_withna$date == date,]$steps)))
    {
        data_withna[is.na(data_withna) & data_withna$date == date,]$steps =                mean(data_withna[data_withna$date == date,]$steps, na.rm = TRUE)
    }
}
```
**Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. **
```{r}
total_steps_per_day_withna = aggregate(data_withna$steps, 
                                by = list(data_withna$date), 
                                sum)$x
total_steps_per_day_withna

```
**Make a histogram of the total number of steps taken each day**
```{r qplot1, fig.width=8, fig.height=6, message=FALSE}
library(ggplot2)
hist(total_steps_per_day_withna, 
     main = "Histogram of Steps/Day", 
     ylab = "Frequency", 
     xlab = "Steps/day")
```

** Do these values differ from the estimates from the first part of the assignment?**
```{r}
summary(total_steps_per_day_withna)
summary(total_steps_per_day)
```
**What is the impact of imputing missing data on the estimates of the total daily number of steps?**

I do not see any visible impact.

## Are there differences in activity patterns between weekdays and weekends?

**Create a new factor variable in the dataset with two levels - "weekday"" and "weekend"" indicating whether a given date is a weekday or weekend day.**
```{r}
weeknames <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
data$day <- factor((weekdays(as.Date(data$date)) %in% weeknames), 
                   levels=c(FALSE, TRUE), 
                   labels=c('weekend', 'weekday'))
```
**Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.**
```{r lattice, fig.width=8, fig.height=6, message=FALSE}
library(lattice)
average_daily_activity_pattern = aggregate(data$steps, 
                                           by = list(data$day, data$interval), 
                                           mean)
with(average_daily_activity_pattern, 
     xyplot(x ~ Group.2|Group.1, 
            type = 'l', 
            xlab = 'Interval',
            ylab = 'Number of Steps',
            layout=(c(1,3))))
```